---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

## Prediction

We will predict the variable "Classe" which represents a gym exercise with 5 different outputs: 
Class A - According to the specification <- Correct execution
Class B - Throwing the elbows to the front <- Common mistake
Class c - Lifting the dumbbell only halfway <- Common mistake
Class D - Lowering the dumbbell only halfway <- Common mistake
Class E - Throwing the hips to the front <- Common mistake

## Resolution Approach
We will create a model with a decision tree and perform cross-validation. We will create two sets: original training data set (75%) and cross validation data set (25%).

```{r}
# Load datasets
trainSet <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"), na.strings = c("NA", "#DIV/0!", ""))
testSet <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"), na.strings = c("NA", "#DIV/0!", ""))

# Remove unused stuff
trainSet <- trainSet[, colSums(is.na(trainSet)) == 0]
trainSet <- trainSet[, -c(1:7)]
testSet <- testSet[, colSums(is.na(testSet)) == 0]
testSet <- testSet[, -c(1:7)]

# Create cross validation set
library(caret)

crossSet <- createDataPartition(trainSet$classe,  p = 0.75, list = F)
trainSetFinal <- trainSet[crossSet, ]
crossSetFinal <- trainSet[-crossSet, ]
```

## Decision Tree

```{r}
library(rpart)

dtModel <- rpart(classe ~ ., data = trainSetFinal, method = "class")
dtPred <- predict(dtModel, crossSetFinal, type = "class")

confusionMatrix(dtPred, as.factor(crossSetFinal$classe))
```

## Random Forest

```{r}
library(randomForest)

rfModel <- randomForest(factor(classe) ~. , data = trainSetFinal, method = "class")
rfPred <- predict(rfModel, crossSetFinal, type = "class")

confusionMatrix(rfPred, as.factor(crossSetFinal$classe))
```

## Selected Option

We pick Random Forest approach because it shows better accuracy (99% vs 74%) and the actual Kappa values (99% vs. 68%) are also much higher. The expected out-of-sample error is estimated less than 1%.

## Results

```{r}
rfPredFinal <- predict(rfModel, testSet, type = "class")

rfPredFinal
```
